---
title: "Topic: Relationship between class level and some cardiovascular indicators of music students."
author: "Jin, Trisha, and Mark"
format: revealjs
editor: visual
---

## Introduction

This study explores the relationship between class level and physiological responses (heart rate and blood pressure) in music students. Using data collected via IoT devices, we analyzed 100 students across three class levels—Beginner, Intermediate, and Advanced—to examine whether more advanced training is linked to higher physiological activity.

## Methodology {#sec-methodology}

- Augmented by IoT devices and AI algorithms.
- Gathers data on student performance, physiological data, engagement, and other metrics in a classroom context.
- Mimics a learning environment in which students' physiological reactions, behavioral patterns, and numerous music performance indicators are tracked by IoT devices during music classes.
- Organized to allow for study of the effects of music instruction.

------------------------------------------------------------------------

- In this research, we will be looking at 3 specific variables: `Class_Level`, and the 2 cardiovascular indicators in the data, which are `Heart_Rate` and `Blood_Pressure`.
- In this data, there are a total of 3 different class levels: beginner, intermediate, and advanced. There are 100 observations in this data.

------------------------------------------------------------------------

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)

music_data = read_csv("music_dataset/music_education_dataset.csv") |>
  # read in the data
  select(Class_Level, Heart_Rate, Blood_Pressure) |>
  # select the 3 columns
  rename("Heart_Rate_(BPM)" = Heart_Rate,
         "Blood_Pressure_(mmHg)" = Blood_Pressure)
  # include units into heart rate and blood pressure
```

```{r fig-class_levels, fig.cap="Student counts of different class levels", echo=FALSE}
class_order = c("Beginner", "Intermediate", "Advanced")
music_data |> ggplot(aes(Class_Level |> factor(levels = class_order))) +
  geom_bar() +
  xlab("Class_Level")
```

------------------------------------------------------------------------

- @fig-class_levels graph shows the comparison of student counts in different class levels.
- There are some differences in the distribution of students in the 3 class levels, but overall, it is still relatively even.
- The students will be grouped by their class levels.

## Result

This result section looks at the relationship between blood pressure, heart rate, and class level among 100 music students in three groups: advanced, intermediate, and beginner.

------------------------------------------------------------------------

#### Averages of heart rate

```{r, echo=FALSE}
ggplot(music_data, aes(x = Class_Level |> factor(levels = class_order),
                       y = `Heart_Rate_(BPM)`)) + 
  geom_boxplot() + 
  labs(title = "Heart Rate by Class Level", y = "Heart Rate (BPM)", x = "Class Level")
```

---

Beginners: around 85 BPM

Intermediate: around 83 BPM

Advanced: around 80 BPM

The Beginner group displays the highest median heart rate, followed by Intermediate, and Advanced students with the lowest. This shows that students in higher class levels have lower heart rates.

------------------------------------------------------------------------

#### Averages of blood pressure

```{r, echo=FALSE}
ggplot(music_data, aes(x = Class_Level |> factor(levels = class_order),
                       y = `Blood_Pressure_(mmHg)`)) +
 geom_boxplot() +
 labs(title = "Blood Pressure by Class Level", y
= "Blood Pressure (mmHg)", x = "Class Level")
```

---

Advanced: around 121 mmHg

Intermediate: around 120 mmHg

Beginner: around 118 mmHg

The Advanced group displays the highest median blood pressure, while Intermediate students follow closely, and Beginners have the lowest. This means students in higher class levels tend to have higher blood pressure.

## Conclusion

The analysis shows an inverse relationship between heart rate and class level, and a direct relationship between blood pressure and class level:

-   **Heart Rate**: Beginners show the highest median heart rate (\~85 BPM), while Advanced students have the lowest (\~80 BPM). 

-   **Blood Pressure**: In contrast, blood pressure tends to increase with class level—Beginners at \~118 mmHg and Advanced students at \~121 mmHg. 

------------------------------------------------------------------------

The results show that beginner students have higher heart rates, while advanced students have higher blood pressure. This may indicate that beginners experience more performance anxiety, whereas advanced students face greater physical or cognitive demands during music instruction.

## Recommendation

1.Collect additional data (e.g., stress levels) to better understand these physiological patterns.

2.Support beginners with stress-reducing strategies.

3.Use physiological feedback to adapt teaching strategies.

4.Ensure breaks and support for advanced students to manage workload.

## Reference

-   R Core Team (2025). *R: A Language and Environment for Statistical Computing*. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.
-   Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” *Journal of Open Source Software*, *4*(43), 1686. doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.

------------------------------------------------------------------------

-   Ziya. “Music Education Performance Data.” Kaggle.com, 2024, <http://www.kaggle.com/datasets/ziya07/music-education-performance-data>.
