---
title: "introduction"
author: "JTM"
format: html
editor: visual
---

## Introduction

Music education not only enhances artistic skills but can also influence physiological and cognitive development in students. With the integration of Internet of Things (IoT) devices and Artificial Intelligence (AI) algorithms, it is now possible to collect and analyze real-time physiological and performance-related data in educational settings.This study explores the relationship between class level and physiological responses (heart rate and blood pressure) in music students. Using data collected via IoT devices, we analyzed 100 students across three class levels—Beginner, Intermediate, and Advanced—to examine whether more advanced training is linked to higher physiological activity.

## Methodology {#sec-methodology}

This dataset, which was augmented by Internet of Things (IoT) devices and Artificial Intelligence (AI) algorithms, was created to assess the efficacy of music education by gathering data on student performance, physiological data, engagement, and other metrics in a classroom context. The data mimics a learning environment in which students' physiological reactions, behavioral patterns, and numerous music performance indicators are tracked by IoT devices during music classes. With an emphasis on skill development, engagement, and performance outcomes, the dataset is organized to allow for study of the effects of music instruction. (["Music Education Performance Data." by Ziya](https://www.kaggle.com/datasets/ziya07/music-education-performance-data))

In this research, we will be looking at 3 specific variables: `Class_Level`, and the 2 cardiovascular indicators in the data, which are `Heart_Rate` and `Blood_Pressure`. For simplicity purposes, we will only be looking at the first dataset, which is the one that has `Blood_Pressure` as one of its variables, as the new version does not have `Blood_Pressure`. In this data, there are a total of 3 different class levels: beginner, intermediate, and advanced. There are 100 observations in this data.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)

music_data = read_csv("music_dataset/music_education_dataset.csv") |>
  # read in the data
  select(Class_Level, Heart_Rate, Blood_Pressure) |>
  # select the 3 columns
  rename("Heart_Rate_(BPM)" = Heart_Rate,
         "Blood_Pressure_(mmHg)" = Blood_Pressure)
  # include units into heart rate and blood pressure
```

```{r fig-class_levels, fig.cap="Student counts of different class levels", echo=FALSE}
music_data |> ggplot(aes(Class_Level)) + geom_bar()
```

@fig-class_levels graph shows the comparison of student counts in different class levels. There are some differences in the distribution of students in the 3 class levels, but overall, it is still relatively even. To analyze this data, the students will be grouped by their class levels, and within those different class levels, we will try to find if there is any pattern or trend when looking at the heart rate and blood pressure of the students.

## Result
This section examines the relationship between blood pressure and heart rate and class level among 100 music students in three different class levels: advanced, intermediate, and beginner.

1. Class Level Distribution: 
The chart below shows the number of students in each class level. The majority are in the Intermediate group, followed by Beginner and Advanced.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
music_data <- read_csv("music_education_dataset
.csv")
```

```{r, echo=FALSE}
music_data |>
 ggplot(aes(Class_Level)) +
 geom_bar() +
 labs(title = "Student Distribution by Class
Level")
```

2. Averages of blood pressure and heart rate:
```{r, echo=FALSE}
ggplot(music_data, aes(x = Class_Level, y =
Heart_Rate)) +
 geom_boxplot() +
 labs(title = "Heart Rate by Class Level", y =
"Heart Rate (BPM)", x = "Class Level")
```
Beginner: Average Heart Rate around 75 BPM

Intermediate: Average Heart Rate around 80 BPM

Advanced: Average Heart Rate around 85 BPM

```{r, echo=FALSE}
ggplot(music_data, aes(x = Class_Level, y =
Blood_Pressure)) +
 geom_boxplot() +
 labs(title = "Blood Pressure by Class Level", y
= "Blood Pressure (mmHg)", x = "Class Level")
```
Beginner: Blood Pressure around 120 mmHg

Intermediate: Blood Pressure around 125 mmHg

Advanced: Blood Pressure around 130 mmHg

3. Trends: 
According to the research, Advanced students have slightly higher Heart Rate and Blood Pressure than Beginners and intermediate students. Both variables show this pattern, which implies that higher class levels are linked to more physiological reactions.

