---
title: "Report"
author: "JTM"
format: html
editor: visual
---

## Introduction

Music education not only enhances artistic skills but can also influence physiological and cognitive development in students. With the integration of Internet of Things (IoT) devices and Artificial Intelligence (AI) algorithms, it is now possible to collect and analyze real-time physiological and performance-related data in educational settings.This study explores the relationship between class level and physiological responses (heart rate and blood pressure) in music students. Using data collected via IoT devices, we analyzed 100 students across three class levels—Beginner, Intermediate, and Advanced—to examine whether more advanced training is linked to higher physiological activity.

## Methodology {#sec-methodology}

This dataset, which was augmented by Internet of Things (IoT) devices and Artificial Intelligence (AI) algorithms, was created to assess the efficacy of music education by gathering data on student performance, physiological data, engagement, and other metrics in a classroom context. The data mimics a learning environment in which students' physiological reactions, behavioral patterns, and numerous music performance indicators are tracked by IoT devices during music classes. With an emphasis on skill development, engagement, and performance outcomes, the dataset is organized to allow for study of the effects of music instruction. (["Music Education Performance Data." by Ziya](https://www.kaggle.com/datasets/ziya07/music-education-performance-data))

In this research, we will be looking at 3 specific variables: `Class_Level`, and the 2 cardiovascular indicators in the data, which are `Heart_Rate` and `Blood_Pressure`. For simplicity purposes, we will only be looking at the first dataset, which is the one that has `Blood_Pressure` as one of its variables, as the new version does not have `Blood_Pressure`. In this data, there are a total of 3 different class levels: beginner, intermediate, and advanced. There are 100 observations in this data.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)

music_data = read_csv("music_dataset/music_education_dataset.csv") |>
  # read in the data
  select(Class_Level, Heart_Rate, Blood_Pressure) |>
  # select the 3 columns
  rename("Heart_Rate_(BPM)" = Heart_Rate,
         "Blood_Pressure_(mmHg)" = Blood_Pressure)
  # include units into heart rate and blood pressure
```

```{r fig-class_levels, fig.cap="Student counts of different class levels", echo=FALSE}
music_data |> ggplot(aes(Class_Level)) + geom_bar()
```

@fig-class_levels graph shows the comparison of student counts in different class levels. There are some differences in the distribution of students in the 3 class levels, but overall, it is still relatively even. To analyze this data, the students will be grouped by their class levels, and within those different class levels, we will try to find if there is any pattern or trend when looking at the heart rate and blood pressure of the students.

## Result
This result section looks at the relationship between blood pressure, heart rate, and class level among 100 music students in three groups: advanced, intermediate, and beginner.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
music_data <- read_csv("music_education_dataset.csv")
```

#### Averages of heart rate 

```{r, echo=FALSE}
ggplot(music_data, aes(x = Class_Level, y = Heart_Rate)) + 
  geom_boxplot() + 
  labs(title = "Heart Rate by Class Level", y = "Heart Rate (BPM)", x = "Class Level")
```
Beginners: around 85 BPM

Intermediate: around 83 BPM

Advanced: around 80 BPM

The Beginner group displays the highest median heart rate, followed by Intermediate, and Advanced students with the lowest. This shows that students in higher class levels have lower heart rates.

#### Averages of blood pressure 
```{r, echo=FALSE}
ggplot(music_data, aes(x = Class_Level, y =
Blood_Pressure)) +
 geom_boxplot() +
 labs(title = "Blood Pressure by Class Level", y
= "Blood Pressure (mmHg)", x = "Class Level")
```
Advanced: around 121 mmHg

Intermediate: around 120 mmHg

Beginner: around 118 mmHg

The Advanced group display the highest median blood pressure, while Intermediate students follow closely, and Beginners have the lowest. This means students in higher class levels tend to have higher blood pressure.

## Conclusion

Results show a clear upward trend in both heart rate and blood pressure with class level:

-   **Heart Rate**: Averages increase from 75 BPM (Beginner) to 85 BPM (Advanced).
-   **Blood Pressure**: Rises from 120 mmHg to 130 mmHg across the same groups.

This suggests that more advanced music training may be more physically or emotionally demanding. However, other factors such as stress, fitness, or individual variability could also influence these readings.

## Recommendation

1.Conduct longitudinal studies to track changes over time.

2.Include stress or emotion-related variables for a fuller picture.

3.Use physiological feedback to adapt teaching strategies.

4.Ensure breaks and support for advanced students to manage workload.
